Formalizing Stack Safety as a Security Property

### 2021-09-06 submission3/response.txt

> I am in favor of accepting this paper; however there are a few questions that I would like to be addressed.

*** Questions Begin ***

> Why did you choose this RISCV specification?
> Why didn't you choose to reuse an ARM specification (as there are some of them written in Coq in the literature)?

We chose RISCV for the sake of the evaluation: the enforcement mechanisms we
test are existing micropolicies defined in terms of a reference monitor
attached to a RISCV machine.

The underlying formalism itself is not dependent on the details of the
architecture, only the testing framework to test that architecture is.
Indeed, even wrt. testing, there is little in the tested micropolicies that
depends on RISCV either, but moving to another architecture such as ARM
would require translating the micropolicy rules to apply to that
architecture and its calling conventions.

@@

> Would it be easy to adapt your work to another ISA specification?

Fairly easy. The formalism itself translates naturally to most ISA
specifications. The largest amount of work is in the testing framework -
modifying the example generator to match the new ISA.

We are currently working on extending the work to a model of a Cheri-style
hardware capability machine, Cerise, and have not encountered major
obstacles.

@@

> How realistic is the RISCV specification? In other words, are there any RISCV vulnerabilities that are not taken into account in your model?

[Timing channels; anything else?]  [BCP: Even if the RISCV model is fairly
complete (though it doesn't include stuff like VM management, hypervisor
instructions, ..., right?), we don't generate test cases for most
instructions -- we're really just testing the protection mechanism for
function calls.]

@@

> More generally, what are the limitations of the presented work? For example, in the model, there seems to be no distinction between values and addresses; is it a limitation?

[BCP: ???]

> Another example is the outside qualifier that characterizes only a whole component. Wouldn't you need a finer-grain access control policy (and a qualifier for partially outside components)?

[I don't understand this question, any insight would be helpful.]
[BCP: Me either.  I wonder if we can ask for clarification...]

@@

> Could you comment on the Coq formalization that is provided as a supplementary material? Was it straightforward to write a Coq program from a C++ program? Moreover, I looked at the Coq development and found some admitted lemmas in it.

Writing Coq program from C++: I believe this one is for Rob
Admitted Lemmas:

[BCP: I wonder whether they are asking about the compilation from our C/C++
pseudo-source-code to RISCV...?]

@@

>    Seems overly complex in comparison to other relational reasoning papers.

> Comments for author

> I found the paper did not connect the section well. Section 4 presents a lot of details for a framework, which is then used in Section 5. Seem very disconnected? For instance, Line 501: ←C\leftarrow_C←C​ this has a different type to the semantics in the previous section. How does it relate? Generally, there are a lot of things that are similar to the previous section, but you are not actually connecting them. Is this because they are different or just missing exposition.

[We should think if there is anything we can suggest here to improve
organization, but it's unclear to me just what the problem was

[BCP: The review here gives one specific example -- what can we say about it?]

@@

> Line 320: defition

[BCP: We should just fix these sorts of things and then not mention them in
the response.]

@@

> Validation is performed using random testing rather than formal proof. However, testing proved sufficient to uncover shortcomings in prior lazy models (which could then be fixed), and mutation testing was used to ensure that the testing approach would catch errors.
> Comments for author

> line 13: variant of our properties and /does so/ correctly, and

> line 94: what about tail calls?

We believe tail calls are a straightforward extension of our framework,
though they are not supported at present.  Since a tail call is both a call
and a return in a single step, our normal handling of returns doesn't
work -- it would interpret the tail call as a call to and immediate return
from a new function activation. We would need to distinguish tail calls from
normal calls and essentially process the call and the return in the reverse
order.

@@

> line 123: "we extend the generation ..." I'm not sure what that means.

> line 210: We /call/ this form...

> line 293: What about exceptions?

Exceptions are another plausible extension. Adding them would involve
interleaving the return target stack with entries that are not return
targets, but "catch targets." Reaching a catch target would pop all return
targets above it, much like returns that violate well-bracketedness already
do.

### 2021-09-06 submission3/response.txt

> I am in favor of accepting this paper; however there are a few questions that I would like to be addressed.

*** Questions Begin ***

> Why did you choose this RISCV specification?
> Why didn't you choose to reuse an ARM specification (as there are some of them written in Coq in the literature)?

We chose RISCV for the sake of the evaluation: the enforcement mechanisms we
test are existing micropolicies defined in terms of a reference monitor
attached to a RISCV machine.
[APT: Is this really true? Nick and Andre's paper uses Alpha.
More accurate to say that we (and Draper, where much of the micropolicies work as been done)
chose RISCV for its simplicity.]

The underlying formalism itself is not dependent on the details of the
architecture, only the testing framework to test that architecture is.
[APT: Is it relevant to point out that we have generic Coq interfaces to
describe the machine and policy?]
Indeed, even wrt. testing, there is little in the tested micropolicies that
depends on RISCV either, but moving to another architecture such as ARM
would require translating the micropolicy rules to apply to that
architecture and its calling conventions.

@@

> Another example is the outside qualifier that characterizes only a whole component. Wouldn't you need a finer-grain access control policy (and a qualifier for partially outside components)?

[I don't understand this question, any insight would be helpful.]
[BCP: Me either.  I wonder if we can ask for clarification...]

> Could you comment on the Coq formalization that is provided as a supplementary material? Was it straightforward to write a Coq program from a C++ program? Moreover, I looked at the Coq development and found some admitted lemmas in it.

Writing Coq program from C++: I believe this one is for Rob
Admitted Lemmas:
[APT: The admits don't seem to be important...]

[BCP: I wonder whether they are asking about the compilation from our C/C++
pseudo-source-code to RISCV...?]
[APT: ...perhaps under the misconception that we are showing RISCV code in
the examples?]

@@

> Other comments

> The supplementary material is not mentioned in the submission. Links to the Coq formalization would have been useful (e.g., for definition 7.4 that I did not fully understand in the submission, as K' is defined but not used, and so is n'').

The supplementary material should indeed be mentioned in the submission.
[Explanation of definition 7.4:]
[APT: Can we include a list of important links here in the response?]

@@

> The first example programs (Fig. 1 to 4) are C programs. The programs in Fig. 5 and 6 are qualified as C++ programs, but they could be C programs and they are close to the first programs. Moreover, the C syntax for passing parameters by reference is surprising. I would have written f(&x)(line 934) and void f(int *a)(line 944).

[BCP: We should check our C / C++ syntax.]
[APT: I think the reviewer is confused. C doesn't really have call-by-reference: (explicitly) passing a pointer
is _not_ the same thing as (implicitly) passing a reference, as the latter cannot be used in a first-class way 
in the callee: that's the whole distinction we're trying to make starting at line 954 of the text.
Evidently we need to clarify this.]

@@

> Line 320: defition

[BCP: We should just fix these sorts of things and then not mention them in
the response.]

> Line 329: I was really interested in address taken at this point, and whether the model could handle it. I really think the informal aspects of the extensions, should be pulled forward, so these questions are answered earlier

[BCP: This is a good expositional idea.]
[APT: Right, so we should acknowledge it politely -- but we did have reasons for ordering
things the way we did....]

### 2021-09-06 submission3/response.txt

> I am in favor of accepting this paper; however there are a few questions that I would like to be addressed.

*** Questions Begin ***

> Why did you choose this RISCV specification?
> Why didn't you choose to reuse an ARM specification (as there are some of them written in Coq in the literature)?

We chose RISCV for the sake of the evaluation: the enforcement mechanisms we
test are existing micropolicies defined in terms of a reference monitor
attached to a RISCV machine.
[APT: Is this really true? Nick and Andre's paper uses Alpha.
More accurate to say that we (and Draper, where much of the micropolicies work as been done)
chose RISCV for its simplicity.]
[BCP: es.]

@@

> More generally, what are the limitations of the presented work? For example, in the model, there seems to be no distinction between values and addresses; is it a limitation?

[BCP: ???]

> Another example is the outside qualifier that characterizes only a whole component. Wouldn't you need a finer-grain access control policy (and a qualifier for partially outside components)?

[I don't understand this question, any insight would be helpful.]
[BCP: Me either.  I wonder if we can ask for clarification...]
[APT: Seems like reviewer is thinking of components as having internal structure and
wondering if we need to label different parts of the structure in different ways...]

### 2021-09-07 submission3/response.txt

> I am in favor of accepting this paper; however there are a few questions that I would like to be addressed.

*** Questions Begin ***

> Why did you choose this RISCV specification?
> Why didn't you choose to reuse an ARM specification (as there are some of them written in Coq in the literature)?

We chose RISCV for the simplicity. It is a prominent base ISA for
the PIPE extensions that implement the tag monitor for the same reason.

The underlying formalism itself is not dependent on the details of the
architecture, as in fact the properties are defined in terms of a generic
interface over machine and micropolicy.

Even wrt. testing, there is little in the tested micropolicies that
depends on RISCV either, but moving to another architecture such as ARM
would require translating the micropolicy rules to apply to that
architecture and its calling conventions.

@@

Fairly easy. The formalism itself translates naturally to most ISA
specifications. The largest amount of work is in the testing framework -
modifying the example generator to match the new ISA.

We are currently working on extending the work to a model of a Cheri-style
hardware capability machine, Cerise, and have not encountered major
obstacles.

> How realistic is the RISCV specification? In other words, are there any RISCV vulnerabilities that are not taken into account in your model?

We use the RISCV specification to demonstrate that our properties are testable in
a realistic setting. We do not exercize the entire RISCV specification, and the validity
of the model is not dependent on the absence of vulnerabilities in the RISCV implementation.
That said, the RISCV specification we used is fairly realistic, although lacking some features
such as VM management instructions.

@@

> More generally, what are the limitations of the presented work? For example, in the model, there seems to be no distinction between values and addresses; is it a limitation?

Currently, some discussions of our model's limitations is scattered throughout the paper.
We will expand and organize it into an explicit discussion, as follows.
Our goal is to specifically characterize stack safety.
So we are particularly concerned with limitations that prevent the model from applying to realistic
uses of the stack, and not with other structures such as the heap, which we intentionally make no
guarantees about.

Features that we do not currently support:
- Dynamic code generation: this would be rather challenging to integrate into
  our framework.
- Address-taken locals: we have discussed why we don't consider our model suitable
  for handling.
- Exceptions, though we can see a way to extend the model to include them:
  essentially, we would model a "setjump" operation by inserting a special
  target in the return target stack. We would weaken well-bracketedness to
  allow reach such a target from any depth. Since integrity and confidentiality
  are distinct from well-bracketedness and already permit the popping of multiple
  functions from the stack, they can simply treat the exception's target as another
  return target.
- Tail calls, but again we can see how we might extend the model for them:
  though they are not supported at present.  Since a tail call is both a call
  and a return in a single step, our normal handling of returns doesn't
  work -- it would interpret the tail call as a call to and immediate return
  from a new function activation. We would need to distinguish tail calls from
  normal calls and essentially process the call and the return in the reverse
  order.

As for the distinction between values and addresses, conventional ISAs do not make this distinction.
Ongoing work on the Cerise capability machine formalization has required some modification to distinguish
values from addresses, as they are not the same in a capability machine, but it has little effect on the 
model itself. Further, distinguishing between values and addresses is a higher-level abstraction that can
"hide" potential bugs resulting from implicit conversions between them, so it may even be preferable to model
them as the same when possible.

@@

> Another example is the outside qualifier that characterizes only a whole component. Wouldn't you need a finer-grain access control policy (and a qualifier for partially outside components)?

There appears to be some confusion arising from the term "components."
In our model, components are already fine-grained: individual registers
and memory addresses. We recognize that this is a poor choice of term
given the common usage of component to refer to a larger module in a
compartmentalization context, and will change the term to [SNA: still
looking for a word. Atom?]
Since components/atoms are so fine-grained, there is no need for finer
access control.

@@

See below.

> Other comments

> The supplementary material is not mentioned in the submission. Links to the Coq formalization would have been useful (e.g., for definition 7.4 that I did not fully understand in the submission, as K' is defined but not used, and so is n'').

The supplementary material should indeed be mentioned in the submission.
The Coq formalization provided contains two parts: the formalization of our properties,
which was a conceptual aid in their development and may be useful to readers' intuitions,
and the testing framework, which is used to validate our properties by showing that
computable versions of them are testable.

Admitted lemmas are the beginnings of an effort to formalize [details]

In the case of definition 7.4, this is a typo. The K in the third bullet should be K', and the
last instance of n' should be n''.

[APT: Can we include a list of important links here in the response?]

@@

>    Seems overly complex in comparison to other relational reasoning papers.

> Comments for author

> I found the paper did not connect the section well. Section 4 presents a lot of details for a framework, which is then used in Section 5. Seem very disconnected? For instance, Line 501: ←C\leftarrow_C←C​ this has a different type to the semantics in the previous section. How does it relate? Generally, there are a lot of things that are similar to the previous section, but you are not actually connecting them. Is this because they are different or just missing exposition.

[We should think if there is anything we can suggest here to improve flow]

### 2021-09-07 submission3/response.txt

>    Seems overly complex in comparison to other relational reasoning papers.

> Comments for author

> I found the paper did not connect the section well. Section 4 presents a lot of details for a framework, which is then used in Section 5. Seem very disconnected? For instance, Line 501: ←C\leftarrow_C←C​ this has a different type to the semantics in the previous section. How does it relate? Generally, there are a lot of things that are similar to the previous section, but you are not actually connecting them. Is this because they are different or just missing exposition.

Many of the content that seems disconnected takes the form of a parameterizable relation and its various
instatiations in different versions of the property. \rightarrow_C parameterized by the type ctx, which is section 5
is instantiated as (K -> SD)x(list(target)). In later sections it will be instantiated by other types.
Then for any \rightarrow_C, we can extend \rightharpoonup to form \rightharpoonup_C.

We can do more to call out which parts of the formalization are parameterized and when they are being
instantiated, so that the connections are more clear.

@@

> Line 320: defition

[BCP: We should just fix these sorts of things and then not mention them in
the response.]

> Line 329: I was really interested in address taken at this point, and whether the model could handle it. I really think the informal aspects of the extensions, should be pulled forward, so these questions are answered earlier

This is a good point, and we will examine ways to introduce that information sooner. However, we are
hesitant to spend too much time on variants before introducing the basic formalism, because it can
be a lot to keep track of.

[BCP: This is a good expositional idea.]
[APT: Right, so we should acknowledge it politely -- but we did have reasons for ordering
things the way we did....]
[SNA: That was roughly the reason, as I recall? I may take a stab tomorrow at sketching what
moving them up would look like.]

### 2021-09-07 submission3/response.txt

The supplementary material should indeed be mentioned in the submission.
The Coq formalization provided contains two parts: the formalization of our properties,
which was a conceptual aid in their development and may be useful to readers' intuitions,
and the testing framework, which is used to validate our properties by showing that
computable versions of them are testable.

Admitted lemmas are the beginnings of an effort to formalize [details]

In the case of definition 7.4, this is a typo. The K in the third bullet should be K', and the
last instance of n' should be n''.

Map of the supplementary material:
Core contains the abstract machine and policy models, as well as the RISCV instantiations with different policies.
Properties contains Coq definitions corresponding to the different Properties:
- Trace.v and ObsTrace.v define the trace model and observation model
- TraceProperties.v defines arbitrary stack properties given a call trace and context type
- SubroutineSimple.v defines the basic stack safety properties
- SubroutineShare.v defines the version enhanced with argument passing on the stack
- Coroutine.v defines the coroutine version
- Lazy.v defines the lazy properties
- CalleeSave.v defines stack safety with callee-save registers
Testing contains the test system, with generators for different machines and the final test run
via extraction in TestProperties.v

### 2021-09-07 submission3/response.txt

> This submission is well-written, with well-explained ideas. Examples of C and assembly programs are used throughout the submission to illustrate the various définitions. The formalization is provided as a supplementary material together with the test cases. The techniques and results of the submission are interesting and general enough to be reused in different settings.

> I am in favor of accepting this paper; however there are a few questions that I would like to be addressed.

*** Questions Begin ***

> Why did you choose this RISCV specification?
> Why didn't you choose to reuse an ARM specification (as there are some of them written in Coq in the literature)?
> Would it be easy to adapt your work to another ISA specification?

We chose RISCV for its simplicity, and because it has been used as the
base ISA for much of the existing work (by others) to design
hardware that supports tag-based micro-policies.

To clarify a possible misunderstanding: the Coq specification of RISCV
that we use was developed by others (see footnote p. 14) and we were
able to adapt it to our testing framework with relatively little
effort.

Our stack safety theory itself is not dependent on the details of the
architecture, as in fact the properties are defined in terms of a
generic interface over machine and enforcement policy; [see ... in the
Coq development]

Performing testing on another architecture such as ARM would require
modifying the testing framework by example generator for the new ISA,
and translating the micropolicy rules to apply to that architecture
and its calling conventions.  This would take some work, but should be
straightforward.

@@

We are currently working on extending the work to a model of a Cheri-style
hardware capability machine, Cerise, and have not encountered major
obstacles.

> How realistic is the RISCV specification? In other words, are there any RISCV vulnerabilities that are not taken into account in your model?

We use the RISCV specification to demonstrate that our properties are testable in
a realistic setting. We do not exercise the entire RISCV specification, and the validity
of the model is not dependent on the absence of vulnerabilities in the RISCV implementation.
That said, the RISCV specification we use is complete for user mode integer instructions
(IV32IM).

### 2021-09-07 submission3/response.txt

> How realistic is the RISCV specification? In other words, are there any RISCV vulnerabilities that are not taken into account in your model?

We use the RISCV specification to demonstrate that our properties are testable in
a realistic setting. We do not exercise the entire RISCV specification, and the validity
of the model is not dependent on the absence of vulnerabilities in the RISCV implementation.
That said, the RISCV specification we use is complete for user mode integer instructions
(IV32IM).

> More generally, what are the limitations of the presented work? For example, in the model, there seems to be no distinction between values and addresses; is it a limitation?

Currently, discussion of our model's limitations is scattered
throughout the paper.  We will expand and organize it into an explicit
section, as follows.

Our goal is to specifically characterize stack safety.  So we are
particularly concerned with limitations that prevent the model from
applying to realistic uses of the stack, and not with other structures
such as the heap, which we intentionally make no guarantees about.

Features that we do not currently support:

- Address-taken locals. As we discuss at line 954ff, a satisfactory
  treatment of these seems to require integration with a more general
  model of memory safety that covers the heap and static globals too.

- Exceptions. We can see a way to extend the model to include them:
  essentially, we would model a "setjump" operation by inserting a
  special target in the return target stack. We would weaken
  well-bracketedness to allow reaching such a target from any
  depth. Since integrity and confidentiality are distinct from
  well-bracketedness and already permit the popping of multiple
  functions from the stack, they can simply treat the exception's
  target as another return target.

- Tail calls. Again, we can see how we might extend the model for them
  Since a tail call is both a call and a return in a single step, our
  normal handling of returns doesn't work -- it would interpret the
  tail call as a call to and immediate return from a new function
  activation. We would need to distinguish tail calls from normal
  calls and essentially process the call and the return in the reverse
  order.

- Dynamic code generation. This would be rather challenging to
  integrate into our framework, because we rely on annotations in the
  code.
[APT: Right? Can we imagine how this might be supported in future?]

- Concurrency. We can handle a limited form of coroutines, as
  described in 8.3. But handling true concurrency would require
  significant changes in our underlying machine model.

@@

As for the distinction between values and addresses, conventional ISAs
do not make this distinction.  Ongoing work on the Cerise capability
machine formalization has required some modification to distinguish
values from addresses, as they are not the same in a capability
machine, but it has little effect on the model itself. Further,
distinguishing between values and addresses is a higher-level
abstraction that can "hide" potential bugs resulting from implicit
conversions between them, so it may even be preferable to model them
as the same when possible.

> Another example is the outside qualifier that characterizes only a whole component. Wouldn't you need a finer-grain access control policy (and a qualifier for partially outside components)?

There may be some confusion arising from our use of the term
"components."  In our model (see line 398), components are already
fine-grained: individual registers and memory addresses. We recognize
that this is a poor choice of term given the common usage of component
to refer to a larger module in a compartmentalization context, and
will change the term to [SNA: still looking for a word. Atom?
APT: "state element" ?]  Since components/atoms are so fine-grained, there
is no need for finer access control.

### 2021-09-07 submission3/response.txt

The supplementary material should indeed be mentioned in the submission.
The Coq formalization provided contains two parts: the formalization of our properties,
which was a conceptual aid in their development and may be useful to readers' intuitions,
and the testing framework, which is used to validate our properties by showing that
computable versions of them are testable.

Admitted lemmas are the beginnings of an effort to formalize [details]

In the case of definition 7.4, this is a typo. The K in the third bullet should be K', and the
last instance of n' should be n''.

Map of the supplementary material:
Core contains the abstract machine and policy models, as well as the RISCV instantiations with different policies.
Properties contains Coq definitions corresponding to the different Properties:
- Trace.v and ObsTrace.v define the trace model and observation model
- TraceProperties.v defines arbitrary stack properties given a call trace and context type
- SubroutineSimple.v defines the basic stack safety properties
- SubroutineShare.v defines the version enhanced with argument passing on the stack
- Coroutine.v defines the coroutine version
- Lazy.v defines the lazy properties
- CalleeSave.v defines stack safety with callee-save registers
Testing contains the test system, with generators for different machines and the final test run
via extraction in TestProperties.v
[APT: Can we give pointers from each paper definition to corresponding Coq definition?]

@@

> The first example programs (Fig. 1 to 4) are C programs. The programs in Fig. 5 and 6 are qualified as C++ programs, but they could be C programs and they are close to the first programs. Moreover, the C syntax for passing parameters by reference is surprising. I would have written f(&x)(line 934) and void f(int *a)(line 944).

Only Fig. 6 uses C++ features.  Although C programmers might indeed
try to simulate call-by-reference by passing a pointer as suggested
here, this is subtly different from a genuine by-reference argument in
C++, because the latter cannot be directly used in a first-class way
in the callee (without using the & operator, as e.g. at line 945).
Thus, legal access to the value of a by-reference argument is
restricted to the immediate callee (and any chain of sub-callees to
which the argument is handed down by reference), and we can capture
this access pattern in our policies.  On the other hand, pointer
arguments might escape into arbitrary sub-callees, and there is no
obvious way for policies in our style to specify when access through
these pointers should be permitted. We will improve the discussion at
line 954ff to clarify this point.

@@

>    Seems overly complex in comparison to other relational reasoning papers.

> Comments for author

> I found the paper did not connect the section well. Section 4 presents a lot of details for a framework, which is then used in Section 5. Seem very disconnected? For instance, Line 501: ←C\leftarrow_C←C​ this has a different type to the semantics in the previous section. How does it relate? Generally, there are a lot of things that are similar to the previous section, but you are not actually connecting them. Is this because they are different or just missing exposition.

Many of the content that seems disconnected takes the form of a parameterizable relation and its various
instantiations in different versions of the property. \rightarrow_C parameterized by the type ctx, which is section 5
is instantiated as (K -> SD)x(list(target)). In later sections it will be instantiated by other types.
Then for any \rightarrow_C, we can extend \rightharpoonup to form \rightharpoonup_C.

We will do more to call out which parts of the formalization are parameterized and when they are being
instantiated, so that the connections are more clear.

@@

[APT: Add pointer]

> line 118: please briefly summarize what "stepwise" means here. Doing so should not need the full formalism, just a notion of steps. Simply hoisting the discussion from line 521 would likely suffice.

> line 123: "we extend the generation ..." I'm not sure what that means.

[APT: Add a clarification that is is "generation-by-execution" ?]

### 2021-09-07 submission3/response.txt

- Address-taken locals. As we discuss at line 954ff, a satisfactory
  treatment of these seems to require integration with a more general
  model of memory safety that covers the heap and static globals too.

- Exceptions. We can see a way to extend the model to include them:
  essentially, we would model a "setjump" operation by inserting a
  special target in the return target stack. We would weaken
  well-bracketedness to allow reaching such a target from any
  depth. Since integrity and confidentiality are distinct from
  well-bracketedness and already permit the popping of multiple
  functions from the stack, they can simply treat the exception's
  target as another return target.

- Tail calls. Again, we can see how we might extend the model for them
  Since a tail call is both a call and a return in a single step, our
  normal handling of returns doesn't work -- it would interpret the
  tail call as a call to and immediate return from a new function
  activation. We would need to distinguish tail calls from normal
  calls and essentially process the call and the return in the reverse
  order.

- Dynamic code generation. This would be rather challenging to
  integrate into our framework, because we rely on annotations in the
  code. We can easily envision supporting dynamic code generation
  where the generated code is guaranteed never to contain a call
  or other annotated instruction. We could also support annotated-code
  generation in which existing code is copied along with its annotations.

@@

- Concurrency. We can handle a limited form of coroutines, as
  described in 8.3. But handling true concurrency would require
  significant changes in our underlying machine model.

As for the distinction between values and addresses, conventional ISAs
do not make this distinction.  Ongoing work on the Cerise capability
machine formalization has required some modification to distinguish
values from addresses, as they are not the same in a capability
machine, but it has little effect on the model itself. Further,
distinguishing between values and addresses is a higher-level
abstraction that can "hide" potential bugs resulting from implicit
conversions between them, so it may even be preferable to model them
as the same when possible.

> Another example is the outside qualifier that characterizes only a whole component. Wouldn't you need a finer-grain access control policy (and a qualifier for partially outside components)?

There may be some confusion arising from our use of the term
"components."  In our model (see line 398), components are already
fine-grained: individual registers and memory addresses. We recognize
that this is a poor choice of term given the common usage of component
to refer to a larger module in a compartmentalization context, and
will change the term to "state elements." Since components/elements
are so fine-grained, there is no need for finer access control.

### 2021-09-08 submission3/response.txt

> Validation is performed using random testing rather than formal proof. However, testing proved sufficient to uncover shortcomings in prior lazy models (which could then be fixed), and mutation testing was used to ensure that the testing approach would catch errors.
> Comments for author

> line 13: variant of our properties and /does so/ correctly, and

> line 94: what about tail calls?

Tail calls are discussed above in our proposed "limitations" section.
They are not currently supported, but we describe how they might be
implemented.

@@

> line 118: please briefly summarize what "stepwise" means here. Doing so should not need the full formalism, just a notion of steps. Simply hoisting the discussion from line 521 would likely suffice.

> line 123: "we extend the generation ..." I'm not sure what that means.

Line 123 should read "... we extend the generation-by-execution techniques...",
referring to the method of randomly generating valid low-level programs by lazily adding
instructions while simulating execution.

@@

> line 210: We /call/ this form...

> line 293: What about exceptions?

Exceptions are also discussed about in our "limitations" section,
along with a description of what it would take to implement them.

### 2021-09-08 submission3/response.txt

> The first example programs (Fig. 1 to 4) are C programs. The programs in Fig. 5 and 6 are qualified as C++ programs, but they could be C programs and they are close to the first programs. Moreover, the C syntax for passing parameters by reference is surprising. I would have written f(&x)(line 934) and void f(int *a)(line 944).

Only Fig. 6 uses C++ features.  Although C programmers might indeed
try to simulate call-by-reference by passing a pointer as suggested
here, this is subtly different from a genuine by-reference argument in
C++, because the latter cannot be directly used in a first-class way
in the callee (without using the & operator, as e.g. at line 945).
Thus, legal access to the value of a by-reference argument is
restricted to the immediate callee (and any chain of sub-callees to
which the argument is handed down by reference), and we can capture
this access pattern in our policies.  On the other hand, pointer
arguments might escape into arbitrary sub-callees, and there is no
obvious way for policies in our style to specify when access through
these pointers should be permitted. We will improve the discussion at
line 954ff to clarify this point.

>    (310, 886,895,947) The return instruction is confusing as the return type of f is void.
>    (568) two many closing parentheses (twice)
>    (796) I was expecting m' in (m,p',ctx').
>    (1005) lack of space between the last two rules

@@

This is a good point, and we will examine ways to introduce that information sooner. However, we are
hesitant to spend too much time on variants before introducing the basic formalism, because it can
be a lot to keep track of.

[BCP: This is a good expositional idea.]
[APT: Right, so we should acknowledge it politely -- but we did have reasons for ordering
things the way we did....]
[SNA: That was roughly the reason, as I recall? I may take a stab tomorrow at sketching what
moving them up would look like.]

> Line 349: How does this related to control flow integrity/guard, and Intel's CET?

> Line 508: "isRet" is a relation, but you are using as a function.
